Fast Hyper-Parameter Selection for Graph Kernels via Subsampling and\\Multiple Kernel Learning

### 2016-11-26 paper.tex
removed half page

@@

\subsection{Multiple Kernel Learning}
\label{MKL}
MKL \cite{Bach2004,Gonen2011} is one of the most popular paradigms used to learn kernels in real world applications \cite{Bucak2014,Castro2014}. % Zien2007,Wang2015a,Sun
The kernels generated by these techniques are combinations of a prescribed set of basic kernels $\KK_1,...,\KK_R$ with a constraint in the form:
$
	H_R^q = \{ \xx \mapsto \textbf{w} \cdot \pphi_\KK(\xx) : \KK = \sum_{r=1}^R \eta_r \KK_r, \mmu \in \Psi_q, \|\textbf{w}\|_2 \leqslant 1 \}
$
with $\Psi_q = \{ \mmu : \mmu \succcurlyeq 0, \| \mmu \|_q = 1 \}$ and considering the function $\pphi_\KK$ as the feature mapping from the input space to the feature space. The value $q$ being the kind of mean used, is typically fixed to $1$ or $2$.

@@

These algorithms are supported by several theoretical results that bound the \emph{estimation error} (i.e. the difference between the true error and the empirical margin error). These bounds exploit the \emph{Rademacher complexity} applied to the combination of kernels \cite{Maurer2012,Srebro2006,Cortes2009c}%,Hussain2011,Hussain2011a}. %,Kakade2009,Micchelli2005,Kloft2012

@@

On the other hand, the \emph{Optimization based} approaches learn the combination parameters by solving a single optimization problem directly integrated in the learning machine %(e.g. structural risk based target function)
or formulated as a different model (e.g. alignment, or other kernel similarity maximization) \cite{Rakotomamonjy2008,Bach2004,Varma2009}. %,Kloft2011,SorenSonnenburg2006,Vishwanathan2010,Xu2010

@@

\subsubsection{EasyMKL}
\label{EasyMKL}
EasyMKL \cite{Aiolli2015} is a recent MKL algorithm able to combine sets of basic kernels by solving a simple quadratic optimization problem. Besides its proved empirical effectiveness, a clear advantage of EasyMKL compared to other MKL methods is its high scalability with respect to the number of kernels to be combined. Specifically, its computational complexity is constant in memory and linear in time.

EasyMKL finds the coefficients $\boldsymbol{\eta}$ that maximize the margin on the training set, where the margin is computed as the distance between the convex hulls of positive and negative examples. In particular, the general problem EasyMKL tries to optimize is the following:
\vspace{-0.4cm}
\begin{equation}
\label{eq:easymklotig}
\max_{\boldsymbol{\eta}:||\boldsymbol{\eta}||_2=1} \min_{\boldsymbol{\gamma} \in \Gamma} \mbox { }  \boldsymbol{\gamma}^{\top} \YY ( \sum_{r=0}^R \eta_r \KK_r) \YY \boldsymbol{\gamma} + \Lambda ||\boldsymbol{\gamma}||^2.
\end{equation}
\vspace{-0.3cm}

where $\YY$ is a diagonal matrix with training labels on the diagonal, and $\Lambda$ is a regularization hyper-parameter. The domain $\Gamma$ represents two probability distributions over the set of positive and negative examples of the training set, that is $\Gamma = \{\boldsymbol{\gamma} \in \Real_+^{\ell} \hspace{0.2cm} |  \sum_{y_i = +1} \boldsymbol{\gamma}_i = 1 , \sum_{y_i = -1} \boldsymbol{\gamma}_i = 1\}.
$
Note that any element $\boldsymbol{\gamma} \in \Gamma$ corresponds to a pair of points, the first in the convex hull of positive training examples and the second in the convex hull of negative training examples.
At the solution, the first term of the objective function represents the obtained margin, that is the (squared) distance between a point in the convex hull of positive examples and a point in the convex hull of negative examples, in the compounded feature space.
%The objective function in Eq. \ref{eq:easymklotig} can be interpreted as the dual problem of a regularized empirical objective function using the kernel $\sum_{r=1}^R \eta_r \KK_r$.
Eq. \ref{eq:easymklotig} is a minimax problem that can be reduced  to a simple quadratic problem with a technical derivation described  in \cite{Aiolli2015}. %The solution of the quadratic problem is an optimal $\ggamma^*$ for the original min-max formulation. Due to the particular structure of EasyMKL, it is sufficient to provide the average kernel of all the basic kernels ($\KK^A = \frac{1}{R} \sum_{r=1}^R \frac{\KK_r}{Tr(\KK_r)}$). From $\ggamma^*$, it is easy to obtain the optimal weights for the single basic kernels $\KK_r$ by using the following formula
%\begin{equation}
%\label{eq:easyeta}
%	\eta_r = \gamma^{*T} \YY \,\, (\KK_r/Tr(\KK_r)) \,\, \YY \gamma^*, \,\,\,\, \forall r=1,\dots,R.
%\end{equation}

In the following sections, we will refer to this method as EasyMKL\footnote{EasyMKL implementation: github.com/jmikko/EasyMKL}.
%and we have to provide it only $\KK^A$, the dataset $\XX$ and $\yy$ to evaluate equation \ref{eq:easyeta}, and the EasyMKL regularization hyper-parameter $\Lambda$.

@@

The kernels used in our experiments are: all NSPDK combinations with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$, all ODDST combinations with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8,0.9,1.0,1.1,1.2,1.4,1.6,1.8\}$, all WL with $h \in \{0,1\dots9\}$.
In all models, we considered kernels calculated in less that 1 hour, rejecting others.
The baselines used to compare our method consist on a SVM with kernels (a)\kval: the kernel is the one which has the best validation score, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: the kernel is a combination of all base kernels using EasyMKL.
For each baseline, the computation of all kernels is required.
The hyperparameters are chosen using a nested 10-fold CV procedure, and they are $C\in\{2^i, i:-3\dots6\}$, and $\Lambda\in\{0,0.1\dots0.4,0.6\dots1\}$ for \kmkl.

>>>>

\renewcommand{\baselinestretch}{.96}

@@

\subsection{Multiple Kernel Learning}
\label{MKL}
MKL \cite{Gonen2011} is one of the most popular paradigms used to learn kernels in real world applications \cite{Bucak2014,Castro2014}. % Zien2007,Wang2015a,Sun
The kernels generated by these techniques are combinations of a prescribed set of basic kernels $\KK_1,...,\KK_R$ with a constraint in the form:
$
	H_R^q = \{ \xx \mapsto \textbf{w} \cdot \pphi_\KK(\xx) : \KK = \sum_{r=1}^R \eta_r \KK_r, \mmu \in \Psi_q, \|\textbf{w}\|_2 \leqslant 1 \}
$
with $\Psi_q = \{ \mmu : \mmu \succcurlyeq 0, \| \mmu \|_q = 1 \}$ and considering the function $\pphi_\KK$ as the feature mapping from the input space to the feature space. The value $q$ being the kind of mean used, is typically fixed to $1$ or $2$.

@@

These algorithms are supported by several theoretical results that bound the \emph{estimation error} (i.e. the difference between the true error and the empirical margin error). These bounds exploit the \emph{Rademacher complexity} applied to the combination of kernels \cite{Maurer2012,Cortes2009c}%,Hussain2011,Hussain2011a}. %,Kakade2009,Micchelli2005,Kloft2012,Srebro2006,

@@

On the other hand, the \emph{Optimization based} approaches learn the combination parameters by solving a single optimization problem directly integrated in the learning machine %(e.g. structural risk based target function)
or formulated as a different model (e.g. alignment, or other kernel similarity maximization) \cite{Rakotomamonjy2008,Varma2009}. %,Kloft2011,SorenSonnenburg2006,Vishwanathan2010,Xu2010,Bach2004

@@

\subsubsection{EasyMKL}
\label{EasyMKL}
EasyMKL \cite{Aiolli2015} is a recent MKL algorithm able to combine sets of basic kernels by solving a simple quadratic optimization problem. Besides its proved empirical effectiveness, a clear advantage of EasyMKL compared to other MKL methods is its high scalability with respect to the number of kernels to be combined. Specifically, its computational complexity is constant in memory and linear in time.
EasyMKL finds the coefficients $\boldsymbol{\eta}$ that maximize the margin on the training set, where the margin is computed as the distance between the convex hulls of positive and negative examples. In particular, the general problem EasyMKL tries to optimize is the following:
\vspace{-0.4cm}
\begin{equation}
\label{eq:easymklotig}
\max_{\boldsymbol{\eta}:||\boldsymbol{\eta}||_2=1} \min_{\boldsymbol{\gamma} \in \Gamma} \mbox { }  \boldsymbol{\gamma}^{\top} \YY ( \sum_{r=0}^R \eta_r \KK_r) \YY \boldsymbol{\gamma} + \Lambda ||\boldsymbol{\gamma}||^2,
\end{equation}
%\vspace{-0.1cm}
where $\YY$ is a diagonal matrix with training labels on the diagonal, and $\Lambda$ is a regularization hyper-parameter. The domain $\Gamma$ represents two probability distributions over the set of positive and negative examples of the training set, that is $\Gamma = \{\boldsymbol{\gamma} \in \Real_+^{\ell} \hspace{0.2cm} |  \sum_{y_i = +1} \boldsymbol{\gamma}_i = 1 , \sum_{y_i = -1} \boldsymbol{\gamma}_i = 1\}.
$
Note that any element $\boldsymbol{\gamma} \in \Gamma$ corresponds to a pair of points, the first in the convex hull of positive training examples and the second in the convex hull of negative training examples.
At the solution, the first term of the objective function represents the obtained margin, that is the (squared) distance between a point in the convex hull of positive examples and a point in the convex hull of negative examples, in the compounded feature space.
%The objective function in Eq. \ref{eq:easymklotig} can be interpreted as the dual problem of a regularized empirical objective function using the kernel $\sum_{r=1}^R \eta_r \KK_r$.
Eq. \ref{eq:easymklotig} is a minimax problem that can be reduced  to a simple quadratic problem with a technical derivation described  in \cite{Aiolli2015}. %The solution of the quadratic problem is an optimal $\ggamma^*$ for the original min-max formulation. Due to the particular structure of EasyMKL, it is sufficient to provide the average kernel of all the basic kernels ($\KK^A = \frac{1}{R} \sum_{r=1}^R \frac{\KK_r}{Tr(\KK_r)}$). From $\ggamma^*$, it is easy to obtain the optimal weights for the single basic kernels $\KK_r$ by using the following formula
%\begin{equation}
%\label{eq:easyeta}
%	\eta_r = \gamma^{*T} \YY \,\, (\KK_r/Tr(\KK_r)) \,\, \YY \gamma^*, \,\,\,\, \forall r=1,\dots,R.
%\end{equation}
In the following sections, we will refer to this method as EasyMKL\footnote{EasyMKL implementation: github.com/jmikko/EasyMKL}.
%and we have to provide it only $\KK^A$, the dataset $\XX$ and $\yy$ to evaluate equation \ref{eq:easyeta}, and the EasyMKL regularization hyper-parameter $\Lambda$.

@@

The kernels used in our experiments are: all NSPDK combinations with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$, all ODDST combinations with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8\dots1.2,1.4,1.6,1.8\}$, all WL with $h \in \{0,1\dots9\}$.
In all models, we considered kernels calculated in less that 1 hour, rejecting others.
The baselines used to compare our method consist on a SVM with kernels (a)\kval: the kernel is the one which has the best validation score, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: the kernel is a combination of all base kernels using EasyMKL.
For each baseline, the computation of all kernels is required.
The hyperparameters are chosen using a nested 10-fold CV procedure, and they are $C\in\{2^i, i:-3\dots6\}$, and $\Lambda\in\{0,0.1\dots1\}$ for \kmkl.
